---
title: "index"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
source("scripts/scatter_plot_index.R")
source("scripts/top_10_scores_2019_piechart.R")
source("scripts/top_10_scores_2021_barchart.R")
source("scripts/summary_info.R")
```

# Global Health Security Around the World

## Introduction
With the world we live in today, global health security plays a big role in how countries are able to operate, all the way from the leader to the normal civilians. We are interested in how to make best use of open datasets to spread awareness of global health security so it can be improved and help public safety. Specifically we are hoping to visualize and identify countries with low security indexes to inform ourselves and eventually develop ways to help improve their benchmarks. To do so, we referred to the *GHS Index* which has been gathering data from 2019 till now, and is responsible for creating index scores for each couturiers health security.


## Summary Information
```{r Summary}
GHS_summary <- read_csv("GHS-Index.csv")

summary <- summary_info(GHS_summary)
summary$total_observations
summary$lowest_Score
summary$highest_Score
summary$avg_Score_2021
```

The dataset that we had used consisted of many different countries around the world in every continent with specific
parameters such as overall index score, prevention of the emergence or release of pathogens, laws involving containing pathogens/diseases, and more. The chosen dataset includes information from the years 2019 and 2021. In total, the GHS Index has calculated the scores of `r summary$total_observations` countries in the years 2019 and 2021, the country with the lowest score was `r summary$lowest_Score` with `r lowest_Score_S`, and the country with the highest score was `r summary$highest_Score` with `r summary$highest_Score_S`. The average index score in 2021 was `r summary$avg_Score_2021`. 


## Table of Aggregated Data
 ```{r table of GHS INDEX in 2019}
 table
 ```

## Charts

### Scatterplot
```{r Scatterplot of Selected Countries in 2019}
scatterplot
```

The scatterplot shows the dots of selected values in 2019 with each country's name written over the datapoint. The scatterplot here shows a wide range of numbers, indicating the contrasting indexes between countries. There also seems to be many countries clustered around each other as well depending on the continent they are in.

### Pie Chart
```{r Pie Chart of Top 10 Countries in 2019}
piechart
```

The pie chart shows the Top 10 Countries in 2019 with the highest index scores. We can see that all of the top 10 countries fall within 10 points of each other. None of the countries stand out from the rest, it is more of a continuous line.

### Bar Chart
```{r Bar Chart of Top 10 Countries in 2021}
barchart
```

The bar chart shows the Top 10 Countries with the highest scores in 2021. Each bar represents the score of the country itself. We can see that only 3 countries had a score over 70 and no country had a score above 80.
